---
title: "RWorksheet_Gagante#4c.Rmd"
author: "Liza Claire Gagante"
date: "2024-11-03"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Use the dataset mpg
a. Show your solutions on how to import a csv file into the environment.
#The simplest way to read a CSV file is by using the read.csv() function. This function will automatically assume that the first row contains headers.
#If your CSV file uses a semicolon (;) as the separator instead of a comma, you can use read.csv2().
#You can also use read.table(), which is more flexible but requires specifying the separator and other parameters.
```{r}
# read.csv()
# read.csv2()
# read.table()
```


b. Which variables from mpg dataset are categorical?
#manufacturer
#model
#trans
#drv
#fl
#class

c. Which are continuous variables?
#displ
#year
#cyl
#cty
#hwy

2. Which manufacturer has the most models in this data set? Which model has the most variations?
Show your answer.
#Toyota has 6 models

a. Group the manufacturers and find the unique models. Show your codes and result.
```{r}
mpg_data <- read.csv("/cloud/project/WORKSHEET4/mpg.csv")

model_counts <- aggregate(model ~ manufacturer, mpg_data, function(x) length(unique(x)))

model_counts
```
b.Graph the result by using plot() and ggplot(). Write the codes and its result.
```{r}
barplot(model_counts$model, 
        names.arg = model_counts$manufacturer, 
        xlab = "Manufacturer", 
        ylab = "Number of Unique Models", 
        main = "Number of Unique Models per Manufacturer", 
        col = "lightblue", 
        las = 2, 
        cex.names = 0.8) 
```
```{r}
library(ggplot2)
ggplot(model_counts, aes(x = reorder(manufacturer, -model), y = model)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  labs(x = "Manufacturer", y = "Number of Unique Models", 
       title = "Number of Unique Models per Manufacturer") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
2.Same dataset will be used. You are going to show the relationship of the modeland the manufacturer.

a. What does ggplot(mpg, aes(model, manufacturer)) + geom_point() show?
#The y-axis represents different car models (model), and the x-axis represents different manufacturers (manufacturer).
b. For you, is it useful? If not, how could you modify the data to make it more informative?
# Since it just displays which manufacturers have which models, the plot produced by ggplot(mpg, aes(model, manufacturer)) + geom_point() is not very helpful. However, because it doesn't show the frequency of models or other characteristics (such as manufacturing years, engine size, or fuel type), it lacks useful insight.
#To illustrate how many distinct models each manufacturer has made, aggregate the data rather than plotting each model separately. The number of distinct models produced by each manufacturer can then be displayed in a bar chart, which facilitates comparison of the range of options provided by each brand.
#You may see how manufacturers' model options have evolved over time by including production year data. To illustrate trends, such as which manufacturers release new models regularly or maintain popular versions for several years, a plot could, for example, indicate which years each manufacturer released new models.
#You might modify the graphic to show the frequency of each model if some of them appear more than once (perhaps in various configurations). The most popular models for each manufacturer can be identified by using point size, color intensity, or an extra label for each model. This will identify which models were more widely used or had different configurations.

3. Plot the model and the year using ggplot(). Use only the top 20 observations. Write the codes and
its results.
```{r}
library(ggplot2)

top_20 <- mpg_data[1:20, ]

ggplot(top_20, aes(x = model, y = year)) +
  geom_point(color = "blue") +
  labs(x = "Model", y = "Year", title = "Top 20 Observations of Model vs. Year") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
4. Using the pipe (%>%), group the model and get the number of cars per model. Show codes and its result
```{r}
library(dplyr)

model_counts <- mpg_data %>%
  group_by(model) %>%
  summarise(car_count = n())

model_counts
```
a. Plot using geom_bar() using the top 20 observations only. The graphs shoudl have a title, labels and
colors. Show code and results.
```{r}
library(dplyr)
library(ggplot2)

model_counts <- mpg_data %>%
  group_by(model) %>%
  summarise(car_count = n()) %>%
  arrange(desc(car_count)) %>%
  slice_head(n = 20)  

ggplot(model_counts, aes(x = reorder(model, -car_count), y = car_count, fill = model)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  labs(
    title = "Top 20 Car Models by Count",
    x = "Car Model",
    y = "Count"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  

```

b. Plot using the geom_bar() + coord_flip() just like what is shown below. Show codes and its result.
```{r}
ggplot(model_counts, aes(x = reorder(model, -car_count), y = car_count, fill = model)) +
  geom_bar(stat = "identity") +
  labs(title = "Top 20 Car Models by Count",
       x = "Car Model",
       y = "Number of Cars") +
  theme_minimal() +
  scale_fill_discrete(name = "Car Model") +
  coord_flip()

```
5. Plot the relationship between cyl - number of cylinders and displ - engine displacement using
geom_point with aesthetic color = engine displacement. Title should be “Relationship between
No. of Cylinders and Engine Displacement”.
a. How would you describe its relationship? Show the codes and its result.
```{r}
ggplot(mpg_data, aes(x = cyl, y = displ, color = displ)) +
  geom_point() +
  labs(title = "Relationship between No. of Cylinders and Engine Displacement",
       x = "Number of Cylinders",
       y = "Engine Displacement") +
  theme_minimal()
```
#In general, engine displacement (displ) tends to increase in tandem with the number of cylinders. Larger engine displacements should be correlated with higher cylinder counts, as we should anticipate.

6. Plot the relationship between displ (engine displacement) and hwy(highway miles per gallon). Mapped
it with a continuous variable you have identified in #1-c. What is its result? Why it produced such
output?
```{r}
ggplot(mpg_data, aes(x = displ, y = hwy, color = cty)) +
  geom_point() +
  labs(title = "Relationship between Engine Displacement and Highway MPG",
       x = "Engine Displacement",
       y = "Highway Miles per Gallon") +
  theme_minimal()
```
#City MPG (cty) is represented by the color gradient. There is typically a negative correlation between displacement and fuel efficiency, with cars with greater engine displacements (displ) having lower highway (hwy) and city (cty) MPG.
#On highways (high displ), larger engines often use more gasoline, resulting in poorer MPG. A trend where displ has a negative correlation with both hwy and cty would be shown by this result.


6. Import the traffic.csv onto your R environment.
a. How many numbers of observation does it have? What are the variables of the traffic dataset the Show your answer.
```{r}
traffic_data <- read.csv("/cloud/project/WORKSHEET4/traffic.csv")

num_observations <- nrow(traffic_data)
num_variables <- names(traffic_data)

cat("Number of Observations:", num_observations, "\n")
cat("Variables in the Dataset:", num_variables, "\n")
```

b. subset the traffic dataset into junctions. What is the R codes and its output?
```{r}
colnames(traffic_data)

head(traffic_data)

junctions <- split(traffic_data, traffic_data$Junction)

junctions
```

c. Plot each junction in a using geom_line(). Show your solution and output.
```{r}
str(traffic_data)
library(ggplot2)
library(gridExtra)

head(traffic_data)
colnames(traffic_data)

cat("Column names in the dataset:\n")
print(colnames(traffic_data))

plots <- lapply(junctions, function(data) {
  ggplot(data, aes(x = DateTime, y = Vehicles, group = Junction)) + 
    geom_line() +
    labs(title = paste("Traffic Volume at Junction", unique(data$Junction)),
         x = "Time",
         y = "Traffic Volume") +
    theme_minimal()
})
grid.arrange(grobs = plots, ncol = 2)
```
7. From alexa_file.xlsx, import it to your environment
a. How many observations does alexa_file has? What about the number of columns? Show your solution
and answer.
```{r}
library(readxl)

alexa_data <- read_excel("/cloud/project/WORKSHEET4/alexa_file.xlsx")

num_observations <- nrow(alexa_data)
num_columns <- ncol(alexa_data)

cat("Number of Observations:", num_observations, "\n")
cat("Number of Columns:", num_columns, "\n")
```
b. group the variations and get the total of each variations. Use dplyr package. Show solution and
answer.
```{r}
library(dplyr)


variation_counts <- alexa_data %>%
  count(variation)

save(variation_counts, file = "variations.RData")

print(variation_counts)
```
c. Plot the variations using the ggplot() function. What did you observe? Complete the details of the graph. Show solution and answer.
#Variants with the highest totals are clearly displayed with the longest bars.
Variations are sorted by count, making it easy to identify the most and least popular variants at a glance.
The rainbow color scheme visually differentiates each variation, but since it lacks a legend (for simplicity), this plot style works best when variations are already clearly labeled.
```{r}
ggplot(variation_counts, aes(x = reorder(variation, -n), y = n, fill = variation)) +
  geom_bar(stat = "identity", color = "pink", show.legend = FALSE) +
  scale_fill_manual(values = rainbow(length(variation_counts$variation))) +
  labs(title = "Product Variants and Totals", 
       x = "Variation", 
       y = "Total") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) +
  coord_flip() 
```
d. Plot a geom_line() with the date and the number of verified reviews. Complete the details of the
graphs. Show your answer and solution.
```{r}
alexa_data$date <- as.Date(alexa_data$date)

review_counts_by_date <- aggregate(verified_reviews ~ date, data = alexa_data, length)


library(ggplot2)

ggplot(review_counts_by_date, aes(x = date, y = verified_reviews)) +
  geom_line(color = "green", size = 1) +
  labs(
    title = "Verified Reviews Over Time",
    x = "Date",
    y = "Number of Verified Reviews"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
    plot.title = element_text(hjust = 0.5)
  )
```
e. Get the relationship of variations and ratings. Which variations got the most highest in rating? Plot a
graph to show its relationship. Show your solution and answer.
```{r}
library(ggplot2)


avg_rating <- aggregate(rating ~ variation, data = alexa_data, mean)
avg_rating <- avg_rating[order(-avg_rating$rating), ]

ggplot(avg_rating, aes(x = reorder(variation, rating), y = rating, fill = variation)) +
  geom_bar(stat = "identity", color = "yellow", show.legend = FALSE) +
  labs(
    title = "Average Rating by Product Variation",
    x = "Variation",
    y = "Average Rating"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) +
  coord_flip()


```