---
title: "RWorksheet_gagante#1"
author: "Liza Claire Gagante"
date: "2024-09-04"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1.
```{r}
age <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27, 22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 42, 53, 41, 51, 35, 24, 33, 41)

```
a. How many data points?
```{r}
length(age)
```

b. Write the code and it's output.
```{r}
age <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27, 22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 42, 53, 41, 51, 35, 24, 33, 41)
print(age)
```
2.Find the reciprocal of the values for age.
```{r}
reciprocal_age <- 1 / age 
library("MASS")
fractions(reciprocal_age)
```

3. Assign also new_age <- c(age, 0, age)
```{r}
new_age <- c(age, 0, age)
new_age
```
What happen to the new_age?
-it prints the age then it prints zero and prints the age again.

4. Sort the values for age.
```{r}
sorted_age <- sort(age)
sorted_age
```

5. 
```{r}
min_age <- min (age)
max_age <- max (age)
min_age
max_age
```

6. 
```{r}
data <- c (2.4, 2.8, 2.1, 2.5, 2.4, 2.2, 2.5, 2.3, 2.5, 2.3, 2.4, 2.7)
```

a.
```{r}
length(data)
```
b. 
```{r}
data <- c (2.4, 2.8, 2.1, 2.5, 2.4, 2.2, 2.5, 2.3, 2.5, 2.3, 2.4, 2.7)
print(data)
```

7. 
```{r}
doubled_data <- data*2
doubled_data
```
8.1
```{r}
hundred <- seq(1, 100)
hundred
```
8.2 
```{r}
num <- seq(20, 60)
num
```
8.3
```{r}
mean(20:60)
```
8.4
```{r}
sum(51:91)
```
8.5
```{r}
thousands <- seq(1:1000)
thousands
```
a.How many data points from 8.1 to 8.4
```{r}
length(hundred) + length(num) + length(mean)+length(sum)
```
 

b.Write the R code and its output from 8.1 to 8.4
```{r}
length(hundred) + length(num) + length(mean) + length(sum)
```


c.For 8.5 find only maximum data points until 10.
```{r}
max(thousands[thousands <-10])
```

9.
```{r}
Filter(function(i) {all(i %% c(3,5,7) != 0 )}, seq(100))
```

10.
```{r}
ten <- seq(100,1)
ten
```
11. 
```{r}
numbers<-1:24
multiples<-numbers[numbers%%3==0 |
numbers%%5==0]
print(multiples)
```
```{r}
sum_multiples<-sum(multiples)
print(sum_multiples)
```
a.136
b.
```{r}
length(ten)+length(numbers)+length(multiples)+length(sum_multiples)
```

12.
It shows an error unexpected '}'

x <-{0 + x + 5 +}


13.
```{r}
score<- c(72, 86, 92, 63, 88, 89, 91, 92, 75, 75, 77)
score[2]
score[3]
```



