---
title: "RWorksheet_gagante#1"
author: "Liza Claire Gagante"
date: "2024-09-04"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1.Set up a vector named age, consisting of 34, 28, 22, 36, 27, 18, 52, 39, 42, 29,
35, 31, 27, 22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 42, 53, 41,
51, 35, 24, 33, 41.
```{r}
age <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27, 22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 42, 53, 41, 51, 35, 24, 33, 41)

```
a. How many data points?
```{r}
length(age)
```

b. Write the code and it's output.
```{r}
age <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27, 22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 42, 53, 41, 51, 35, 24, 33, 41)
print(age)
```
2.Find the reciprocal of the values for age.
```{r}
reciprocal_age <- 1 / age 
library("MASS")
fractions(reciprocal_age)
```

3. Assign also new_age <- c(age, 0, age)
```{r}
new_age <- c(age, 0, age)
new_age
```
What happen to the new_age?
-it prints the age then it prints zero and prints the age again.

4. Sort the values for age.
```{r}
sorted_age <- sort(age)
sorted_age
```

5. Find the minimum and maximum value for age.
```{r}
min_age <- min (age)
max_age <- max (age)
min_age
max_age
```

6. Set up a vector named data, consisting of 2.4, 2.8, 2.1, 2.5, 2.4, 2.2, 2.5,
2.3, 2.5, 2.3, 2.4, and 2.7.
```{r}
data <- c (2.4, 2.8, 2.1, 2.5, 2.4, 2.2, 2.5, 2.3, 2.5, 2.3, 2.4, 2.7)
```

a.How many data points?
```{r}
length(data)
```
b. Write the R code and its output.
```{r}
data <- c (2.4, 2.8, 2.1, 2.5, 2.4, 2.2, 2.5, 2.3, 2.5, 2.3, 2.4, 2.7)
print(data)
```

7. Generates a new vector for data where you double every value of the data. | What
happen to the data?
```{r}
doubled_data <- data*2
doubled_data
```
8.1 Integers from 1 to 100.
```{r}
hundred <- seq(1, 100)
hundred
```
8.2 Numbers from 20 to 60.
```{r}
num <- seq(20, 60)
num
```
8.3 Mean of numbers from 20 to 60.
```{r}
mean(20:60)
```
8.4 Sum of numbers from 51 to 91
```{r}
sum(51:91)
```
8.5 Integers from 1 to 1,000
```{r}
thousands <- seq(1:1000)
thousands
```
a.How many data points from 8.1 to 8.4
```{r}
length(hundred) + length(num) + length(mean)+length(sum)
```
 

b.Write the R code and its output from 8.1 to 8.4
```{r}
length(hundred) + length(num) + length(mean) + length(sum)
```


c.For 8.5 find only maximum data points until 10.
```{r}
max(thousands[thousands <-10])
```

9. *Print a vector with the integers between 1 and 100 that are not divisible by 3, 5 and
7 using filter option.
```{r}
Filter(function(i) {all(i %% c(3,5,7) != 0 )}, seq(100))
```

10. Generate a sequence backwards of the integers from 1 to 100.
```{r}
ten <- seq(100,1)
ten
```
11. List all the natural numbers below 25 that are multiples of 3 or 5.
Find the sum of these multiples.
```{r}
numbers<-1:24
multiples<-numbers[numbers%%3==0 |
numbers%%5==0]
print(multiples)
```
```{r}
sum_multiples<-sum(multiples)
print(sum_multiples)
```
a. How many data points from 10 to 11?
136
b. Write the R code and its output from 10 and 11.
```{r}
length(ten)+length(numbers)+length(multiples)+length(sum_multiples)
```

12.Statements can be grouped together using braces ‘{’ and ‘}’. A group of statements
is sometimes called a block. Single statements are evaluated when a new line is typed
at the end of the syntactically complete statement. Blocks are not evaluated until a
new line is entered after the closing brace.
It shows an error unexpected '}'

```{r}
#x <-{0 + x + 5 +}
```


13. *Set up a vector named score, consisting of 72, 86, 92, 63, 88, 89, 91, 92, 75,
75 and 77. To access individual elements of an atomic vector, one generally uses the
x[i] construction.
Find x[2] and x[3]. Write the R code and its output.
```{r}
score<- c(72, 86, 92, 63, 88, 89, 91, 92, 75, 75, 77)
score[2]
score[3]
```

14.*Create a vector a = c(1,2,NA,4,NA,6,7).
```{r}
a= c(1,2,NA,4,NA,6,7)
```
a.Change the NA to 999 using the codes print(a,na.print="-999").
```{r}
print(a,na.print="-999")
```
b. Write the R code and its output. Describe the output.
The output is [1]    1    2 -999    4 -999    6    7
In the output, NA values are displayed as -999 as specified by the na.print argument in the print function.

15.A special type of function calls can appear on the left hand side of the assignment
operator as in > class(x) <- "foo".
```{r}
name = readline(prompt="Input your name: ")
age = readline (prompt="Input your age: ")
print(paste("My name is", name, "and I am", age, "years old."))
print(R.version.string)
```

What is the output of the above code?
It prints the sentence that incorporates the user's input.
