---
title: "RWorksheet_Gagante#4b.Rmd"
author: "Liza Claire Gagante"
date: "2024-10-29"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Using Loop Function
for() loop
1. Using the for loop, create an R script that will display a 5x5 matrix as shown in
Figure 1. It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix.
Hint: Use abs() function to get the absolute value
```{r}
vectorA <- c(1, 2, 3, 4, 5)
matrix_5x5 <- matrix( nrow = 5, ncol = 5)

for (i in 1:5) {
  for (j in 1:5) {
    matrix_5x5[i, j] <- abs(i - j)
  }
}

print(matrix_5x5)
```
2. Print the string "*" using for() function. The output should be the same as shown
in Figure
```{r}
for (i in 1:5) {
  line <- rep('"*"', i) 
  cat(line, sep = " ")  
  cat("\n")
}
```
3. Get an input from the user to print the Fibonacci sequence starting from the 1st input
up to 500. Use repeat and break statements. Write the R Scripts and its output.
```{r}
# start <- as.integer(readline(prompt = "Enter the starting number: "))

#a <- start
#b <- 1
#cat(a, b, sep = " ")

#repeat {
 # next_term <- a + b
 # if (!is.na(next_term) && next_term > 500) {
 #   break
#  }
#  cat(next_term, " ")
 # a <- b
#  b <- next_term
  #}
#cat("\n")
```
4. Import the dataset as shown in Figure 1 you have created previously.
a. What is the R script for importing an excel or a csv file? Display the first 6 rows of
the dataset? Show your codes and its result.
```{r}
data <- read.csv("Shoe_sizes.csv")

head(data)
```
b. Create a subset for gender(female and male). How many observations are there in
Male? How about in Female? Write the R scripts and its output.
```{r}
male_data <- subset(data, Gender == "M")
female_data <- subset(data, Gender == "F")

num_males <- nrow(male_data)
num_females <- nrow(female_data)

num_males
num_females
```
c. Create a graph for the number of males and females for Household Data. Use plot(),
chart type = barplot. Make sure to place title, legends, and colors. Write the R scripts and
its result.
```{r}
gender_counts <- table(data$Gender)

barplot(gender_counts,
        main = "Number of Males and Females",
        xlab = "Gender",
        ylab = "Count",
        col = c("blue", "pink"),
        legend = rownames(gender_counts))
```



