---
title: "RWorksheet_gagante#4a.Rmd"
author: "Liza Claire Gagante"
date: "2024-10-17"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. The table below shows the data about shoe size and height. Create a data frame.
```{r}
data <- data.frame(
  Shoe_size = c(6.5, 9.0, 8.5, 8.5, 10.5, 7.0, 9.5, 9.0, 13.0, 7.5, 10.5, 8.5, 12.0, 
                10.5, 13.0, 11.5, 8.5, 5.0, 10.0, 6.5, 7.5, 8.5, 10.5, 8.5, 10.5, 11.0, 9.0),
  Height = c(66.0, 68.0, 64.5, 65.0, 70.0, 64.0, 70.0, 71.0, 72.0, 64.0, 74.5, 71.0, 71.0, 
             77.0, 72.0, 59.0, 62.0, 72.0, 66.0, 64.0, 67.0, 73.0, 69.0, 72.0, 70.0, 69.0, 70.0),
  Gender = c("F", "F", "F", "F", "F", "F", "F", "M", "F", "M", "M", "M", "M", 
             "M", "M", "F", "F", "M", "F", "F", "F", "M", "F", "M", "M", "M", "M")
)
data
```

a. Describe the data.
  
  The table shows data regarding the shoe size, height, and gender of each individuals.

b. Create a subset by males and females with their corresponding shoe size and height.
What its result? Show the R scripts.
```{r}
male_data <- subset(data, Gender == "M", select = c(Shoe_size, Height))

male_data
```

```{r}
female_data <- subset(data, Gender == "F", select = c(Shoe_size, Height))

female_data
```


c. Find the mean of shoe size and height of the respondents. Write the R scripts and its
result.
```{r}
mean_shoe_size <- mean(data$Shoe_size)

mean_shoe_size

mean_height <- mean(data$Height)

mean_height
```

d. Is there a relationship between shoe size and height? Why?
-Yes, beacuse as the height increases, shoe size tends to increase as well.

2. Construct character vector months to a factor with factor() and assign the result to
factor_months_vector. Print out factor_months_vector and assert that R prints out
the factor levels below the actual values.
Consider data consisting of the names of months:
"March","April","January","November","January",
"September","October","September","November","August",
"January","November","November","February","May","August",
"July","December","August","August","September","November","February",|
April")

```{r}
months_vector <- c("March","April","January","November","January",
"September","October","September","November","August",
"January","November","November","February","May","August",
"July","December","August","August","September","November","February",
"April")

factor_months_vector <- factor(months_vector)

print(factor_months_vector)

levels(factor_months_vector)
```
3. Then check the summary() of the months_vector and factor_months_vector. | Interpret the results of both vectors. Are they both equally useful in this case?
```{r}
summary(months_vector)
```
```{r}
summary(factor_months_vector)
```
The result of months_vector it states the Length, Class and Mode. While the factor_months_vector states how many months in the data for example December, December has 1. As what I saw they are both useful because it is easy for me to understand and determine how many types of data from the raw data itself.

4. Create a vector and factor for the table below.
```{r}
direction_vector <- c("East", "West", "North", "West", "North", "West", "North", "West")

factor_data <- factor(direction_vector)

new_order_data <- factor(factor_data,levels = c("East","West","North"))
print(new_order_data)
```
5. Enter the data below in Excel with file name = import_march.csv
a. Import the excel file into the Environment Pane using read.table() function.
Write the code.
```{r}
data_excel <- read.table("import_march.csv")
```

b. View the dataset. Write the R scripts and its result.
```{r}
data_excel
```
6. Full Search
a. Create an R Program that allows the User to randomly select numbers from 1 to 50.Then display the chosen number. If the number is beyond the range of the selected choice,it will have to display a string “The number selected is beyond the range of 1 to 50”. If number 20 is inputted by the User, it will have to display “TRUE”, otherwise display the input number.

```{r}
exhaustive_search <- function(selected_number) {
  
  if(selected_number < 1 || selected_number > 50) {
    print("The selected number is beyond the range of 1 to 50")
  } else if(selected_number == 20) {
    print("TRUE")
  } else {
    print(selected_number)
  }
}
selected_number <- readline(prompt = "Select a number from 1 to 50: ")
  exhaustive_search(selected_number)
```
7. Change
a. Write a function that prints the minimum number of bills that must be paid, given the price of the snack.
Input: Price of snack (a random number divisible by 50) Output: Minimum number of bills needed to purchase a snack.
```{r}
min_bills <- function(price) {
  bills <- c(1000, 500, 200, 100, 50)
  
  num_bills <- 0
  for(bill in bills) {
    count <- price %/% bill
    num_bills <- num_bills + count
    
    price <- price %% bill
  }
  
  print(paste("Minimum number of bills needed to purchase a snack: ", num_bills))
}

min_bills(1650)
```
8. The following is each student’s math score for one semester. Based on this, answer the
following questions.
a. Create a dataframe from the above table. Write the R codes and its output.
```{r}
students <- data.frame(
  Name = c("Annie", "Thea", "Steve", "Hanna"),
  Grade1 = c(85, 65, 75, 95),
  Grade2 = c(65, 75, 55, 75),
  Grade3 = c(85, 90, 80, 100),
  Grade4 = c(100, 90, 85, 90)
)

students
```
b. Without using the rowMean function, output the average score of students whose average
math score over 90 points during the semester. write R code and its output.
Example Output: Annie’s average grade this semester is 88.75.
```{r}
for (i in 1:nrow(students)) {
  total_score <- students$Grade1[i] + students$Grade2[i] + students$Grade3[i] + students$Grade4[i]
  avg_score <- total_score / 4
  
  if (avg_score > 90) {
     formatted_output <- sprintf("%s's average grade this semester is %.2f.", students$Name[i], avg_score)
    print(formatted_output)
  }
}
```
c. Without using the mean function, output as follows for the tests in which the average
score was less than 80 out of 4 tests.
Example output: The nth test was difficult.
```{r}
for (j in 2:5) { 
  total_test_score <- sum(students[, j])
  avg_test_score <- total_test_score / nrow(students)
  
  if (avg_test_score < 80) {
    print(paste("The", j-1, "th test was difficult."))
  }
}
```
d. Without using the max function, output as follows for students whose highest score for
a semester exceeds 90 points.
Example Output: Annie’s highest grade this semester is 95.
```{r}
for (i in 1:nrow(students)) {
  
  grades <- c(students$Grade1[i], students$Grade2[i], students$Grade3[i], students$Grade4[i])
  
  highest_grade <- grades[1]
  for (grade in grades) {
    if (grade > highest_grade) {
      highest_grade <- grade
    }
  }
  
  if (highest_grade > 90) {
    print(paste(students$Name[i], "'s highest grade this semester is", highest_grade))
  }
}
```




