---
title: "RWorksheet_gagante#4a.Rmd"
author: "Liza Claire Gagante"
date: "2024-10-17"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. The table below shows the data about shoe size and height. Create a data frame.
```{r}
shoe_size <- c(6.5, 9.0, 8.5, 8.5, 7.0, 9.0, 9.5, 13.0, 7.5, 10.5, 10.5, 12.0, 10.5, 
               13.0, 11.5, 8.5, 5.0, 10.0, 6.5, 7.5, 8.5, 10.5, 8.5, 10.5, 11.0, 9.0, 13.0)
height <- c(66.0, 68.0, 65.0, 65.0, 64.0, 71.0, 72.0, 72.0, 74.5, 67.0, 74.5, 71.0, 71.0,
            77.0, 72.0, 59.0, 62.0, 72.0, 66.0, 64.0, 67.0, 73.0, 69.0, 72.0, 70.0, 69.0, 70.0)
gender <- c("F", "F", "F", "F", "F", "F", "F", "M", "F", "M", "M", "M", "M", 
            "M", "M", "F", "F", "M", "F", "F", "F", "M", "F", "M", "M", "M", "M")

data <- data.frame(Shoe_size = shoe_size, Height = height, Gender = gender)

data
```

a. Describe the data.
  The table shows data regarding the shoe size, height, and gender of each individuals.

b. Create a subset by males and females with their corresponding shoe size and height.
What its result? Show the R scripts.
```{r}
male_data <- subset(data, Gender == "M", select = c(Shoe_size, Height))

male_data
```

```{r}
female_data <- subset(data, Gender == "F", select = c(Shoe_size, Height))

female_data
```


c. Find the mean of shoe size and height of the respondents. Write the R scripts and its
result.
```{r}
mean_shoe_size <- mean(data$Shoe_size)

mean_shoe_size

mean_height <- mean(data$Height)

mean_height
```

d. Is there a relationship between shoe size and height? Why?


2. Construct character vector months to a factor with factor() and assign the result to
factor_months_vector. Print out factor_months_vector and assert that R prints out
the factor levels below the actual values.
Consider data consisting of the names of months:
"March","April","January","November","January",
"September","October","September","November","August",
"January","November","November","February","May","August",
"July","December","August","August","September","November","February",|
April")

```{r}
months_vector <- c("March","April","January","November","January",
"September","October","September","November","August",
"January","November","November","February","May","August",
"July","December","August","August","September","November","February",
"April")

factor_months_vector <- factor(months_vector)

print(factor_months_vector)

levels(factor_months_vector)
```
3. Then check the summary() of the months_vector and factor_months_vector. | Interpret the results of both vectors. Are they both equally useful in this case?
```{r}
summary(months_vector)
```
```{r}
summary(factor_months_vector)
```
The result of months_vector it states the Length, Class and Mode. While the factor_months_vector states how many months in the data for example December, December has 1. As what I saw they are both useful because it is easy for me to understand and determine how many types of data from the raw data itself.

4. Create a vector and factor for the table below.
```{r}
direction_vector <- c("East", "West", "North", "West", "North", "West", "North", "West")

factor_data <- factor(direction_vector)

new_order_data <- factor(factor_data,levels = c("East","West","North"))
print(new_order_data)
```
5. Enter the data below in Excel with file name = import_march.csv
a. Import the excel file into the Environment Pane using read.table() function.
Write the code.

b. View the dataset. Write the R scripts and its result.
