---
title: "RWorksheet_Gagante#2.R."
author: "Liza Claire Gagante"
date: "2024-09-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Using Vectors
1. Create a vector using : operator
a. Sequence from -5 to 5. Write the R code and its output.
Describe its output.
```{r}
x <- -5:5
x
```
The output is a vector of integers from -5 to 5, inclusive.

b. x <- 1:7. What will be the value of x?
```{r}
x <- 1:7
x
```
The value of x is a vector containing integers from 1 to 7.

2.* Create a vector using seq() function
a. seq(1, 3, by=0.2) # specify step size
Write the R script and its output. Describe the output.
```{r}
x <- seq(1,3, by=0.2)
x
```
This output shows a sequence starting from 1 to 3 with increments of 0.2.

3. A factory has a census of its workers. There are 50 workers in total. The following
list shows their ages: 34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27,
22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 35,
24,33, 41, 53, 40, 18, 44, 38, 41, 48, 27, 39, 19, 30, 61, 54, 58, 26,
18.

```{r}
ages <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27,
22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 35,
24,33, 41, 53, 40, 18, 44, 38, 41, 48, 27, 39, 19, 30, 61, 54, 58, 26,
18)
ages
```

a. Access 3rd element, what is the value?
```{r}
ages[3]
```
22

b. Access 2nd and 4th element, what are the values?
```{r}
ages[c(2,4)]
```
28 and 36

c. Access all but the 4th and 12th element is not
included. Write the R script and its output.

```{r}
ages[-c(4,12)]
```
Output: [1] 34 28 22 27 18 52 39 42 29 35 27 22 37 34 19 20 57 49 50 37 46 25 17 37
[25] 43 53 41 51 35 24 33 41 53 40 18 44 38 41 48 27 39 19 30 61 54 58 26 18

4. *Create a vector x <- c("first"=3, "second"=0, "third"=9). Then named the
vector, names(x).
```{r}
x <- c("first" =3, "second" =0, "third" =9)
```

a. Print the results. Then access x[c("first", "third")].
Describe the output.
```{r}
x
x[c("first","third")]
```
The named vector is accessed for "first" and "third," returning their values.

b. Write the code and its output.
```{r}
x <- c("first" =3, "second" =0, "third" =9)
x
x[c("first","third")]
```
Output: first second  third 
     3      0      9 
first third 
    3     9 
    
5. Create a sequence x from -3:2.
```{r}
x <- -3:2
x
```

a. Modify 2nd element and change it to 0;
x[2] <- 0
x
Describe the output.
```{r}
x[2] = 0
x
```
The second element is modified to 0.

b. Write the code and its output.
```{r}
x <- -3:2
x[2] <- 0
x
```
Output: [1] -3  0 -1  0  1  2

6. *The following data shows the diesel fuel purchased by Mr. Cruz.
Month Jan Feb March Apr May June
Price per liter (PhP) 52.50 57.25 60.00 65.00 74.25 54.00
Purchase–quantity(Liters) 25 30 40 50 10 45
a. Create a data frame for month, price per liter (php)
and purchase-quantity (liter). Write the R scripts and
its output.
```{r}
month <- c("Jan","Feb", "March", "Apr", "May", "June")
price_per_liter <- c(52.50, 57.25, 60.00, 65.00, 74.25, 54.00)
purchase_quantity <- c(25, 30, 40, 50, 10, 45)
fuel_data <- data.frame(month, price_per_liter, purchase_quantity)
fuel_data
```
Its output is a data frame containing month, price per liter, and purchase quantity.

b. What is the average fuel expenditure of Mr. Cruz from
Jan to June? Note: Use ‘weighted.mean(liter,
purchase)‘. Write the R scripts and its output.
```{r}
average_fuel_expenditure <- weighted.mean(price_per_liter, purchase_quantity)
average_fuel_expenditure
```
Output: 59.2625

7. R has actually lots of built-in datasets. For example, the rivers data “gives the lengths
(in miles) of 141 “major” rivers in North America, as compiled by the US Geological
Survey”.
a. Type “rivers” in your R console.
Create a vector data with 7
elements, containing the number of elements (length)
in rivers, their sum (sum), mean (mean),
median(median), variance(var), standard deviation(sd),
minimum (min) and maximum (max).
```{r}
data (rivers)

data <- c(length(rivers), #Number of elements
          sum(rivers),    #Sum of all elements
          mean(rivers),   #Mean
          median(rivers), #Median
          var(rivers),    #Variance
          sd(rivers),     #Standard deviation
          min(rivers),    #Minimum value
          max(rivers)    #Maximum value
          )
data

```

b. What are the results?
141.0000  83357.0000    591.1844    425.0000 243908.4086    493.8708
    135.0000   3710.0000
Length: 141.0000
Sum: 83357.0000
Mean: 591.1844 
Median: 425.0000
Variance: 243908.4086
Standard:  493.8708
Minimum:  135.0000
Maximum: 3710.0000

c. Write the R scripts and its outputs.
```{r}
data (rivers)

data <- c(length(rivers), #Number of elements
          sum(rivers),    #Sum of all elements
          mean(rivers),   #Mean
          median(rivers), #Median
          var(rivers),    #Variance
          sd(rivers),     #Standard deviation
          min(rivers),    #Minimum value
          max(rivers)    #Maximum value
          )
data

```
Output: 
141.0000  83357.0000    591.1844    425.0000 243908.4086    493.8708
    135.0000   3710.0000
Length: 141.0000
Sum: 83357.0000
Mean: 591.1844 
Median: 425.0000
Variance: 243908.4086
Standard:  493.8708
Minimum:  135.0000
Maximum: 3710.0000

8. a. Create vectors according to the above table.
Write the R scripts and its output.
```{r}
power_ranking <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25)
celebrity_name <- c("Tom Cruise", "Rolling Stones", "Oprah Winfrey", "U2", "Tiger Woods", 
                    "Steven Spielberg", "Howard Stern", "50 Cent", "Cast of the Sopranos", 
                    "Dan Brown", "Bruce Springsteen", "Donald Trump", "Muhammad Ali", 
                    "Paul McCartney", "George Lucas", "Elton John", "David Letterman", 
                    "Phil Mickelson", "J.K Rowling", "Bradd Pitt", "Peter Jackson", 
                    "Dr. Phil McGraw", "Jay Lenon", "Celine Dion", "Kobe Bryant")
pay <- c(67, 90, 225, 110, 90, 332, 302, 41, 52, 88, 55, 44, 55, 40, 233, 34, 40, 47, 75, 25, 39, 45, 32, 40, 31)

power_ranking
celebrity_name
pay
```

b. Modify the power ranking and pay of J.K. Rowling.
Change power ranking to 15 and pay to 90. Write the
R scripts and its output.
```{r}
celebrity_name[19] <- "J.K Rowling"
power_ranking[19] <- 15
pay[19] <- 90

power_ranking
celebrity_name
pay
```
Output:
[1]  1  2  3  4  5  6  7  8  9 10 11 12 13 14 15 16 17 18 15 20 21 22 23
[24] 24 25
 [1] "Tom Cruise"           "Rolling Stones"       "Oprah Winfrey"       
 [4] "U2"                   "Tiger Woods"          "Steven Spielberg"    
 [7] "Howard Stern"         "50 Cent"              "Cast of the Sopranos"
[10] "Dan Brown"            "Bruce Springsteen"    "Donald Trump"        
[13] "Muhammad Ali"         "Paul McCartney"       "George Lucas"        
[16] "Elton John"           "David Letterman"      "Phil Mickelson"      
[19] "J.K Rowling"          "Bradd Pitt"           "Peter Jackson"       
[22] "Dr. Phil McGraw"      "Jay Lenon"            "Celine Dion"         
[25] "Kobe Bryant"         
 [1]  67  90 225 110  90 332 302  41  52  88  55  44  55  40 233  34  40
[18]  47  90  25  39  45  32  40  31

c. Create an excel file from the table above and save it
as csv file(PowerRanking). Import the csv file into
the RStudio. What is the R script?
```{r}
imported_data <- read.csv("PowerRanking.csv")
print(imported_data)
```

d. Access the rows 10 to 20 and save it as Ranks.RData.
Write the R script and its output.
```{r}
ranks <- imported_data[10:20, ]

save(ranks, file = "Ranks.RData")
print(ranks)
```

e. Describe its output.
The output is table with rows 10 through 20 in it. This contains the names, corresponding pay, and ranking of the celebrities (from 10 to 20), all of which are kept in a csv file.

9. Download the Hotels-Vienna https://tinyurl.com/Hotels-Vienna
a. Import the excel file into your RStudio.
What is the R script?
```{r}
library(openxlsx)
hotels_vienna <- read.xlsx("hotels-vienna.xlsx")
```
b. How many dimensions does the dataset have?
What is the R script? WHat is its output?
```{r}
dim(hotels_vienna)
```
Output: 428  24

c. Select columns country, neighbourhood,
price, stars, accomodation_type, and
ratings. Write the R script.
```{r}
selcol <- hotels_vienna[, c("country", "neighbourhood", "price", "stars", "accommodation_type", "rating")]
```

d. Save the data as **new.RData to your RStudio. Write the
R script.
```{r}
save(selcol, file = "new.RData")
```

e. Display the first six rows and last six rows of the
new.RData. What is the R script?
```{r}
load("new.RData")
head(selcol)
tail(selcol)
```

10. Create a list of ten (10) vegetables you ate during your lifetime. If none, just list down.
a. Write the R scripts and its output.
```{r}
veg <- list("Lettuce","Cabbage","Carrot","Corn","Cucumber","Potatoes","Pumpkin","Eggplant","Spinach","Tomatoes")
veg
```

b. Add 2 additional vegetables after the last vegetables
in the list. What is the R script and its output?
```{r}
veg <- append(veg,list("Okra","Papaya"))
veg
```
Output:
Lettuce
Cabbage
Carrot
Corn
Cucumber
Potatoes
Pumpkin
Eggplant
Spinach
Tomatoes
Okra
Papaya

c. Add 4 additional vegetables after index 5. How many
datapoints does your vegetable list have? What is the
R script and its output?
```{r}
veg <- append(veg, list("Mushroom","Radish","Green Peas","Asparagus"), after = 5)
veg
```
Output:
Lettuce
Cabbage
Carrot
Corn
Cucumber
Mushroom
Radish
Green Peas
Asparagus
Potatoes
Pumpkin
Eggplant
Spinach
Tomatoes
Okra
Papaya

d. Remove the vegetables in index 5, 10, and 15. How
many vegetables were left? Write the codes and its
output.
```{r}
veg <- veg[-c(5, 10, 15)]
veg
```
Output:
Lettuce
Cabbage
Carrot
Corn
Mushroom
Radish
Green Peas
Asparagus
Pumpkin
Eggplant
Spinach
Tomatoes
Papaya



